Modeling the Circumnuclear Mediums and Accetion Rates of Quiescent Supermassive Black Holes

### 2015-01-27 ms.tex
append

Thermal stability ($\dot{q}_{\rm heat} > \dot{q}_{\rm cool}$; $t_{\rm cool} > t_{\rm ff}$) at the stagnation radius thus requires a critical minimum heating rate
\begin{align}
v_{w} > v_{\rm crit} \simeq
  {\rm max}\begin{cases}
   200 \eta_{0.02}^{0.14} M_{\bullet,8}^{0.11}\,{\rm km\,s^{-1}}  &, \text{core}\\
   200 \eta_{0.02}^{0.17} M_{\bullet,8}^{0.08}\,{\rm km\,s^{-1}}   &, \text{cusp}\\
\sigma,     
  \end{cases}
  \label{eq:cooling2}
\end{align}
Our numerical results in $\S\ref{sec:numerical}$ will demonstrate that $\tilde{v}_{w} \gtrsim v_{\rm crit} \sim 200-300$ km s$^{-1}$ is generally a condition for stability at all radii, not just at the stagnation radius (Fig.~\ref{fig:cooling}).  The requirement that $v_{\rm w} \gtrsim \sigma$ in equation (\ref{eq:cooling2}) results because the top two lines are derived under the assumption that the stagnation radius is given by equation (\ref{eq:rs_simple}).  When $v_{\rm w} \lesssim \sigma$ the stagnation radius moves to larger radii than predicted by by equation (\ref{eq:rs_simple}) or may not exist at all, resulting in much higher gas densities and inevitable cooling instability.

@@

Type Ia supernovae (SNe Ia) represent a potentially important source of heating, which unlike core collapse SNe is present even in an evolved stellar population.  If each Ia supernova injects thermal energy $E_{\rm Ia}$ into the interstellar medium, and the SNe rate (per stellar mass) is given by $R_{\rm   Ia}$, then the resulting volumetric heating rate $E_{\rm Ia}R_{\rm  Ia}$ produces an effective wind heating parameter (equation~[\ref{eq:vw_eff}]) \begin{align} v_{w}^{\rm Ia} =\sqrt{\frac{2 t_h R_{\rm Ia}
E_{\rm Ia}}{\eta}} \label{eq:vw_sne}.
\end{align} The thermal energy injected by a Ia SN $E_{\rm Ia} \simeq
\epsilon_{\rm Ia} 10^{51}$ ergs depends on the efficiency $\epsilon_{\rm Ia}$
with which the initial blast wave energy is converted into bulk or turbulent
motion instead of being lost to radiation.  \cite{Thornton+98}
estimate a radiative efficiency $\epsilon_{\rm Ia} \sim 0.1$,
depending weakly on surrounding density, but \citet{Sharma+14} argues
that $\epsilon_{\rm Ia}$ can be considerably higher, $\sim 0.4$, if
the SNe occur in a hot dilute medium, as may characterize the circumnuclear medium.  We hereafter adopt $\epsilon_{\rm Ia} = 0.4$.

@@

The high value of $v_{w}^{\rm Ia}$ implies that Ia SNe represent an important source of CNM heating.  However, SNe can only be approximated as supplying constant heating throughout space and time if the rate of SNe is rapid compared to the inflow time at the radius of interest (\citealt{ShcherbakovWong+:2014a}).  We define the ``Ia radius"
  \begin{align}
    r_{\rm Ia} \sim \left(\frac{G}{R_{\rm Ia}\sigma}\right)^{1/2} \sim
    38 M_{\bullet,8}^{-0.1}(\tau_{\star}/t_{\rm h})^{0.56}\,{\rm pc}
    \label{eq:rIa}
  \end{align}
as the location exterior of which the time interval between subsequent supernovae $\tau_{\rm Ia} \sim 1/M_{\rm enc}R_{\rm Ia} \sim G/r\sigma^{2}R_{\rm Ia}$ exceeds the local dynamical timescale $t_{\rm
dyn} \sim r/\sigma$, where we again adopt the Ia rate for an old stellar population and in the final equality we estimate the velocity dispersion using the $M_{\bullet}-\sigma$ relationship
(equation~[\ref{eq:Msigma}]).

Assuming that $v_{w}^{\rm Ia} \gg \sigma$ then by substituting $v_{w}^{\rm Ia}$ (eq.~\ref{eq:vIa}) into our expression for the stagnation radius (\ref{eq:rs_simple}), we can write
\begin{eqnarray}
\left.\frac{r_{\rm Ia}}{r_{\rm s}}\right|_{\rm v_{w}^{\rm Ia}} &\approx& 12.4\eta_{0.02} M_{\bullet,8}^{-1.1}(\epsilon_{\rm Ia}/0.4)^{-1}(\tau_{\star}/t_{\rm h})^{1.68} \nonumber \\
&\approx& 12.4M_{\bullet,8}^{-1.1}(\epsilon_{\rm Ia}/0.4)^{-1}(\tau_{\star}/t_{\rm h})^{0.38}
\end{eqnarray}
The fact that $r_{\rm Ia} \gg r_{\rm s}$ would appear to imply that Ia heating only represents a sources of steady heating near the stagnation radius for very massive SMBHs $M_{\bullet} \gtrsim 10^9 \Msun$ and/or very young populations ($\tau_{\star} \ll t_{\rm h}$).  However, this expression underestimates the importance of Ia heating, since the actual stagnation radius will be substantially larger than its value for $v_{w}^{\rm Ia}$ if the nominal heating rate is much less than this.  In general it is reasonable to expect that the impact of Ia SN will increase for larger SMBH masses.

What is the effect of the Ia radius?  Between subsequent supernovae, stars release a gaseous mass $M_{\rm g} \approx \eta M_{\star}\tau_{\rm Ia}/t_{\rm h}$ interior to the Ia radius, which is gravitationally bound to the SMBH by an energy $E_{\rm bind} \approx M_{\rm g}\sigma^{2}$.  Thus, according to the above definitions, $E_{\rm Ia}/E_{\rm bind} \sim (v_{w}^{\rm Ia})^{2}/2\sigma^{2} \gtrsim 1$.  Hence Ia SN can in principle dynamically clear out gas from radii $\sim r_{\rm Ia} \gtrsim r_{\rm s}$.  Hence, even if $r_{\rm s} > r_{\rm Ia}$, the accretion rate will be limited to a value set by substituting $r_{\rm Ia}$ for $r_{\rm s}$ in the derivation of equation $\ref{eq:eddr_analytic}$, i.e.  
\begin{eqnarray}
\left.\frac{\dot{M}}{\dot{M}_{\rm edd}}\right|_{\rm max} &\approx& \frac{\eta M_{\bullet}}{\dot{M}_{\rm edd} t_{\rm h}}\left(\frac{R_{\rm Ia}}{R_{\rm soi}}\right)^{2-\Gamma} \approx \nonumber \\
 && \begin{cases}
    3 \times 10^{-4} M_{\bullet,8}^{-1.26}(\tau_{\star}/t_{\rm h})^{-0.3}
   & \text{core} \\
    1.5 \times 10^{-4} \Mbheight^{-0.77}(\tau_{\star}/t_{\rm h})^{-0.7}   & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_Ia}
\end{eqnarray}

\subsection{Millisecond Pulsars}
 Energy injection from millisecond pulsars (MSPs) is a potentially
important heating source.  If the number of MSPs per unit stellar mass
is $n_{\rm msp}$ and each contributes on average a spin-down
luminosity $\bar{L}_{\rm sd}$, then the resulting heating per unit
volume $e \approx \bar{L}_{\rm sd}n_{\rm msp}\epsilon$ results in an
effective heating rate (eq.~ [\ref{eq:vw_eff}])

\begin{eqnarray} v_{w}^{\rm MSP} \approx
30\left(\frac{\epsilon}{0.1}\right)^{1/2}\left(\frac{\bar{L}_{\rm
sd}}{10^{34}\,{\rm erg\,s^{-1}}}\right)^{1/2} \eta_{0.02}^{-1/2}\,{\rm
km\,s^{-1}},
 \label{eq:vmsp}
  \end{eqnarray}

where $\epsilon$ is the thermalization efficiency of
the wind, normalized to a value $\epsilon \lesssim 0.1$ inferred based
on modeling the insterstellar mediums of globular clusters
\citep{NaimanSoares-Furtado+:2013a}.  The above estimate adopts a
pulsar density $n_{\rm msp}\simeq3 \times 10^{-40} $ MSPs g$^{-1}$
motivated by the $\sim 30,000$ estimated MSPs in the Milky Way of
stellar mass $6\times 10^{10}M_{\odot}$ (\citealt{Lorimer13}).

The average spin-down luminosity of millisecond pulsars in the field
is estimated to be $\bar{L}_{\rm sd} \sim 10^{34}$ erg s$^{-1}$ based
on the ATNF radio pulsar catalog (\citealt{Manchester+05}), in which
case $v_{w} \lesssim 30$ km s$^{-1}$.  For higher spin-down
luminosities, $L_{\rm sd}\simeq 10^{35}$ ergs s$^{-1}$ characteristic
of some Fermi-detected pulsars, then the higher value of $v_{w}
\lesssim 300$ km s$^{-1}$ makes MSP heating in principle important
under the optimistc assumption $\epsilon = 1$.  On the other hand, the
binaries giving rise to MSPs may be disocciated by stellar
interactions in the dense nuclear cluster, reducing their numbers as
compared to the field population estimate above.  MSPs are thus unlikely to be an important source of heating in comparison to stellar heating.

@@

Feedback from accretion onto the SMBH represents an important source of heating which is, however, the most difficult to quantify.  A key difference between AGN heating and other sources discussed thus far is its likely dependence on the mass accretion rate $\dot{M}$, which is itself a function of the heating rate (eq.~[\ref{eq:mdot_analytic}]).

There are two types of SMBH feedback: kinetic and radiative.  Radiative feedback is potentially effective even in low luminosity AGN via Compton heating (e.g.,~\citealt{Gan+14}), which provides a volumetric heating rate (\citealt{Gan+14})
\be
\dot{e}_{\rm C} = 4.1\times 10^{-35}n^{2}\xi T_{\rm C}\,{\rm erg\,cm^{-3}\,s^{-1}},
\ee
where $\xi = L/n r^{2}$ is the ionization parameter and $L$ is the SMBH luminosity with Compton temperature $T_{\rm C} \sim 10^{9}$ K $\gg T$.

{\bf AG: Note that I obtain Compton heating which is a factor of 2
  higher--although power law dependences look right to me. Note that
  the Compton heating will depend on radius--I use rs in my own
  estimate. }
We estimate the importance of Compton heating by assuming the SMBH
radiates with a luminosity $L = \epsilon \dot{M}c^{2}$, where
$\epsilon$ is the radiative efficiency and where $\dot{M}$ is
estimated from equation (\ref{eq:mdot_analytic}).  Then using
equations (\ref{eq:rs_simple}), (\ref{eq:mdot_analytic}),
(\ref{eq:rhors}), (\ref{eq:rhostarrs}) we calculate that from equation
(\ref{eq:vw_eff}) that
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 13 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.38}v_{500}^{-1.4}\,{\rm
km\,s^{-1}} &, \text{core}\\ 20 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.24}v_{500}^{-0.7}\,{\rm
km\,s^{-1}} &, \text{cusp},
  \end{cases}
  \label{eq:vC}
\end{align} where $T_{C,9} = T_{C}/10^{9}$ K and $\epsilon_{-2} =
\epsilon/0.01 \sim 1$ is the maximum radiative efficiency for a thick
accretion flow (e.g.~\citealt{Narayan&Yi95}).  Assuming that heating
is dominated by Compton heating ($\tilde{v}_{w} = v_{w}^{C}$) one finds
that
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 106 \eta_{0.02}^{0.21}T_{\rm
C,9}^{0.21}\epsilon_{-2}^{0.21} M_{\bullet,8}^{0.16}\,{\rm km\,s^{-1}}
&, \text{core}\\ 76 \eta_{0.02}^{0.29}T_{\rm
C,9}^{0.29}\epsilon_{-2}^{0.29} M_{\bullet,8}^{0.14}\,{\rm km\,s^{-1}}
&, \text{cusp},
  \end{cases}
  \label{eq:vC2}
\end{align} Unless $\eta$ is very high, we conclude that Compton
heating is insufficient to stabilize the steady flow against cooling
instability (e.g. $v_{w}^{C} < v_{\rm critt}$;
eq.~[\ref{eq:cooling2}]).

@@

If cooling becomes important in the flow ($\tilde{v}_{w} < v_{\rm crit}$), it may lead to thermal instabilities.  There are two kinds of ``instabilities", global and local.  The global instability is simply that, if cooling becomes very rapid, the gas loses thermal pressure and starts to inflow faster, i.e. the entire CNM accretes.  In
practice this should could manifest as a large increase in the stagnation radius, due an effective {\it decrease} in the effective heating rate.  This global ``instability" is not a true instability insofar as the gas does not necessarily clump locally into low-T clouds.

>>>>

Thermal stability ($\dot{q}_{\rm heat} > \dot{q}_{\rm cool}$; $t_{\rm cool} > t_{\rm ff}$) at the stagnation radius thus requires a critical minimum heating rate
\begin{align}
v_{w} > v_{\rm CI} \simeq
  {\rm max}\begin{cases}
   200 \eta_{0.02}^{0.14} M_{\bullet,8}^{0.11}\,{\rm km\,s^{-1}}  &, \text{core}\\
   200 \eta_{0.02}^{0.17} M_{\bullet,8}^{0.08}\,{\rm km\,s^{-1}}   &, \text{cusp}\\
\sigma,     
  \end{cases}
  \label{eq:cooling2}
\end{align}
Our numerical results in $\S\ref{sec:numerical}$ will demonstrate that $\tilde{v}_{w} \gtrsim v_{\rm CI} \sim 200-300$ km s$^{-1}$ is generally a condition for stability at all radii, not just at the stagnation radius (Fig.~\ref{fig:cooling}).  The requirement that $v_{\rm w} \gtrsim \sigma$ in equation (\ref{eq:cooling2}) results because the top two lines are derived under the assumption that the stagnation radius is given by equation (\ref{eq:rs_simple}).  When $v_{\rm w} \lesssim \sigma$ the stagnation radius moves to larger radii than predicted by by equation (\ref{eq:rs_simple}) or may not exist at all, resulting in much higher gas densities and inevitable cooling instability.

@@

Type Ia supernovae (SNe Ia) represent a potentially important source of heating, which unlike core collapse SNe is present even in an evolved stellar population.  If each Ia supernova injects thermal energy $E_{\rm Ia}$ into the interstellar medium, and the SNe rate (per stellar mass) is given by $R_{\rm   Ia}$, then the resulting volumetric heating rate $E_{\rm Ia}R_{\rm  Ia}$ produces an effective wind heating parameter (equation~[\ref{eq:vw_eff}]) \begin{align} v_{w}^{\rm Ia} =\sqrt{\frac{2 t_h R_{\rm Ia}
E_{\rm Ia}}{\eta}} \label{eq:vw_sne}.
\end{align} The thermal energy injected by a Ia SN $E_{\rm Ia} \simeq
\epsilon_{\rm Ia} 10^{51}$ ergs depends on the efficiency $\epsilon_{\rm Ia}$
with which the initial blast wave energy is converted into bulk or turbulent
motion instead of being lost to radiation.  \cite{Thornton+98}
estimate a radiative efficiency $\epsilon_{\rm Ia} \sim 0.1$,
depending weakly on surrounding density, but \citet{Sharma+14} argues
that $\epsilon_{\rm Ia}$ can be considerably higher, $\sim 0.4$, if
the SNe occur in a hot dilute medium, as may characterize the circumnuclear medium.  Hereafter we adopt $\epsilon_{\rm Ia} = 0.4$ as fiducial.

@@

The high value of $v_{w}^{\rm Ia}$ implies that Ia SNe represent an important source of CNM heating.  However, SNe can only be approximated as supplying constant heating throughout space and time if the rate of SNe is rapid compared to the inflow time at the radius of interest (\citealt{ShcherbakovWong+:2014a}).  We define the ``Ia radius"
  \begin{align}
    r_{\rm Ia} \sim \left(\frac{G}{R_{\rm Ia}\sigma}\right)^{1/2} \sim
    38 M_{\bullet,8}^{-0.1}(\tau_{\star}/t_{\rm h})^{0.56}\,{\rm pc}
    \label{eq:rIa}
  \end{align}
as the location exterior of which the time interval between subsequent supernovae $\tau_{\rm Ia} \sim (M_{\rm enc}R_{\rm Ia})^{-1} \sim G/(r\sigma^{2}R_{\rm Ia})$ exceeds the local dynamical timescale $t_{\rm
dyn} \sim r/\sigma$, where we again adopt the Ia rate for an old stellar population and in the final equality we estimate the velocity dispersion using the $M_{\bullet}-\sigma$ relationship
(equation~[\ref{eq:Msigma}]).

Assuming that $v_{w}^{\rm Ia} \gg \sigma$ then by substituting $v_{w}^{\rm Ia}$ (eq.~\ref{eq:vIa}) into equation (\ref{eq:rs_simple}) for the stagnation radius, we find that
\begin{eqnarray}
\left.\frac{r_{\rm Ia}}{r_{\rm s}}\right|_{\rm v_{w}^{\rm Ia}} &\approx& 12.4\eta_{0.02} M_{\bullet,8}^{-1.1}(\epsilon_{\rm Ia}/0.4)^{-1}(\tau_{\star}/t_{\rm h})^{1.68} \nonumber \\
&\approx& 12.4M_{\bullet,8}^{-1.1}(\epsilon_{\rm Ia}/0.4)^{-1}(\tau_{\star}/t_{\rm h})^{0.38}
\label{eq:rIars}
\end{eqnarray}
The fact that $r_{\rm Ia} \gg r_{\rm s}$ appears to imply that Ia heating can only be approximated as a steady heating source near the stagnation radius for extremely massive SMBHs $M_{\bullet} \gtrsim 10^9 \Msun$ or very young populations ($\tau_{\star} \ll t_{\rm h}$).  However, equation (\ref{eq:rIars}) may underestimate the importance of Ia heating because $r_{\rm Ia}/r_{\rm s}$ will be substantially smaller than its value estimated using $v_{w}^{\rm Ia}$ if the nominal heating rate (without Ia contribution) is much less than this.  Nevertheless, the impact of Ia SN is clearly greater for larger SMBH masses for which $r_{\rm Ia} \lesssim r_{\rm s}$.

What is the effect of the Ia radius?  Between subsequent supernovae, stars release a gaseous mass $M_{\rm g} \approx \eta M_{\star}\tau_{\rm Ia}/t_{\rm h}$ interior to the Ia radius, which is gravitationally bound to the SMBH by an energy $E_{\rm bind} \sim M_{\rm g}\sigma^{2}$.  Thus, according to the above definitions, $E_{\rm Ia}/E_{\rm bind} \sim (v_{w}^{\rm Ia})^{2}/2\sigma^{2} \gtrsim 1$.  Hence Ia SN can in principle dynamically clear out gas from radii $\sim r_{\rm Ia} \gtrsim r_{\rm s}$.  Thus, even in cases when $r_{\rm s} > r_{\rm Ia}$, the SMBH accretion rate is limited to the value set by substituting the Ia radius $r_{\rm Ia}$ (eq.~[\ref{eq:rIa}]) for $r_{\rm s}$ in the derivation of equation $\ref{eq:eddr_analytic}$, i.e.  
\begin{eqnarray}
\left.\frac{\dot{M}}{\dot{M}_{\rm edd}}\right|_{\rm max} &\approx& \frac{\eta M_{\bullet}}{\dot{M}_{\rm edd} t_{\rm h}}\left(\frac{R_{\rm Ia}}{R_{\rm soi}}\right)^{2-\Gamma} \approx \nonumber \\
 && \begin{cases}
    3 \times 10^{-4} M_{\bullet,8}^{-1.26}(\tau_{\star}/t_{\rm h})^{-0.3}
   & \text{core} \\
    1.5 \times 10^{-4} \Mbheight^{-0.77}(\tau_{\star}/t_{\rm h})^{-0.7}   & \text{cusp}.
  \end{cases}
  \label{eq:eddr_Ia}
\end{eqnarray}
{\bf BDM: check!}

\subsection{Millisecond Pulsars}
 Energy injection from millisecond pulsars (MSPs) is a potentially
important heating source.  If the number of MSPs per unit stellar mass
is $n_{\rm msp}$ and each contributes on average a spin-down
luminosity $\bar{L}_{\rm sd}$, then the resulting heating per unit
volume $e \approx \bar{L}_{\rm sd}n_{\rm msp}\epsilon_{\rm msp}$ results in an
effective heating rate (eq.~ [\ref{eq:vw_eff}])
\begin{eqnarray} v_{w}^{\rm MSP} \approx
30\left(\frac{\epsilon_{\rm msp}}{0.1}\right)^{1/2}\left(\frac{\bar{L}_{\rm
sd}}{10^{34}\,{\rm erg\,s^{-1}}}\right)^{1/2} \eta_{0.02}^{-1/2}\,{\rm
km\,s^{-1}},
 \label{eq:vmsp}
  \end{eqnarray} 
where $\epsilon_{\rm msp}$ is the thermalization efficiency of
the wind, normalized to a value $\epsilon \lesssim 0.1$ inferred based
on modeling the insterstellar mediums of globular clusters
\citep{NaimanSoares-Furtado+:2013a}.  The above estimate adopts a
pulsar density $n_{\rm msp}\simeq3 \times 10^{-40} $ MSPs g$^{-1}$
motivated by the $\sim 30,000$ estimated MSPs in the Milky Way of
stellar mass $6\times 10^{10}M_{\odot}$ (\citealt{Lorimer13}).

The average spin-down luminosity of millisecond pulsars in the field
is estimated to be $\bar{L}_{\rm sd} \sim 10^{34}$ erg s$^{-1}$ based
on the ATNF radio pulsar catalog (\citealt{Manchester+05}), in which
case $v_{w} \lesssim 30$ km s$^{-1}$.  For higher spin-down
luminosities, $L_{\rm sd}\simeq 10^{35}$ ergs s$^{-1}$ characteristic
of some Fermi-detected pulsars, then the higher value of $v_{w}
\lesssim 300$ km s$^{-1}$ makes MSP heating in principle important
under the optimistc assumption $\epsilon_{\rm msp} = 1$.  On the other hand, the
binaries giving rise to MSPs may be disocciated by stellar
interactions in the dense nuclear cluster, reducing their numbers as
compared to the field population estimate above.  MSPs are thus unlikely to be an important source of heating in comparison to stellar heating.

@@

Feedback from accretion onto the SMBH represents an important source of heating which, however, is also the most difficult to quantify.  A key difference between AGN heating and the other sources discussed thus far is its dependence on the SMBH accretion rate $\dot{M}$, which is itself a function of the heating rate (eq.~[\ref{eq:mdot_analytic}]).

There are two types of SMBH feedback: kinetic and radiative.  Radiative feedback is potentially effective even in low luminosity AGN via Compton heating, which provides a volumetric heating rate (\citealt{Gan+14})
\be
\dot{e}_{\rm C} = 4.1\times 10^{-35}n^{2}\xi T_{\rm C}\,{\rm erg\,cm^{-3}\,s^{-1}},
\ee
where $\xi = L/n r^{2}$ is the ionization parameter and $L$ is the SMBH luminosity with Compton temperature $T_{\rm C} \sim 10^{9}$ K $\gg T$ (\citealt{Gan+14}).

{\bf AG: Note that I obtain Compton heating which is a factor of 2
  higher--although power law dependences look right to me. Note that
  the Compton heating will depend on radius--I use rs in my own
  estimate.  BDM: if so, the please update the numbers!  }
The importance of Compton heating can be estimated by assuming the SMBH
radiates with a luminosity $L = \epsilon_{\rm rad} \dot{M}c^{2}$, where
$\epsilon_{\rm rad}$ is the radiative efficiency and where $\dot{M}$ is
estimated from equation (\ref{eq:mdot_analytic}).  Then using
equations (\ref{eq:rs_simple}), (\ref{eq:mdot_analytic}),
(\ref{eq:rhors}), (\ref{eq:rhostarrs}) we calculate that from equation
(\ref{eq:vw_eff}) that the effective Compton heating rate at the stagnation radius is given by
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 13 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.38}v_{500}^{-1.4}\,{\rm
km\,s^{-1}} &, \text{core}\\ 20 \eta_{0.02}^{0.5}T_{\rm
C,9}^{0.5}\epsilon_{-2}^{0.5} M_{\bullet,8}^{0.24}v_{500}^{-0.7}\,{\rm
km\,s^{-1}} &, \text{cusp},
  \end{cases}
  \label{eq:vC}
\end{align} where $T_{C,9} = T_{C}/10^{9}$ K and $\epsilon_{-2} =
\epsilon_{\rm rad}/0.01 \sim 1$ is the maximum radiative efficiency for a thick
accretion flow (e.g.~\citealt{Narayan&Yi95}).  If heating
were dominated by Compton heating, i.e. $\tilde{v}_{w} = v_{w}^{C}$, then equation (\ref{eq:vC}) can be solved for $\tilde{v}_{\rm w}$ to find
\begin{align} v_{w}^{\rm C} \simeq
  \begin{cases} 106 \eta_{0.02}^{0.21}T_{\rm
C,9}^{0.21}\epsilon_{-2}^{0.21} M_{\bullet,8}^{0.16}\,{\rm km\,s^{-1}}
&, \text{core}\\ 76 \eta_{0.02}^{0.29}T_{\rm
C,9}^{0.29}\epsilon_{-2}^{0.29} M_{\bullet,8}^{0.14}\,{\rm km\,s^{-1}}
&, \text{cusp},
  \end{cases}
  \label{eq:vC2}
\end{align} 
The fact that $v_{w}^{\rm C} \lesssim 200$ km s$^{-1}$ for $\eta \lesssim 1$ shows that Compton heating alone is insufficient to stabilize the flow against cooling
instability, i.e. $v_{w}^{C} < v_{\rm CI}$ (eq.~[\ref{eq:cooling2}]).

@@

If cooling becomes important in the flow ($\tilde{v}_{w} < v_{\rm CI}$), it may lead to thermal instabilities.  There are two kinds of ``instabilities", global and local.  The global instability is simply that, if cooling becomes very rapid, the gas loses thermal pressure and starts to inflow faster, i.e. the entire CNM accretes.  In
practice this should could manifest as a large increase in the stagnation radius, due an effective {\it decrease} in the effective heating rate.  This global ``instability" is not a true instability insofar as the gas does not necessarily clump locally into low-T clouds.

### 2015-01-27 windheat.tex
editing2
append

The energy and mass injection from stellar winds will be the sum ofthe
contributions from main sequence and post-main sequence (PMS) stars.
For an impulsively formed stellar population of age $t$, the mass
injection rate per unit stellar mass,$\dot{\bar{m}}(t)$, and  the energy
injection rate per unit stellar mass, $\dot{\bar{e}}(t)$,  will be given by

\begin{align} 
  \dot{\bar{m}}(t) &= \frac{\Delta M(t) \mu(M_{\rm TO}(t))
    \left|\dot{M}_{\rm TO}(t)\right| + f_{\rm MS} \int_{m_0}^{m_{\rm
        T}(t)}
    \dot{m}(\Mstar, t) \mu(\Mstar) {\rm d}\Mstar }{\bar{m}_*}\\
  \dot{\bar{e}}(t) &=\dot{e}_{\rm TO}(t)+ f_{\rm MS} \int_{m_0}^{m_{\rm T}(t)}
  \frac{\vw^2(\Mstar, t) \dot{m}(\Mstar, t) \mu(\Mstar) {\rm d}\Mstar}{\bar{m}_*}.
  \label{eq:edotImp}
\end{align}

The first terms in each expression above correspond to the
contributions from PMS stars, while the second terms correspond to the
contributions from main sequence stars. The main sequence winds are a
small fraction of the mass, and they may be not be thermalized and
mixed with the rest of the injected gas. Thus, we include a
thermalization efficiency, $ 0\le f_{\rm MS}\le 1$, in
equation~\eqref{eq:edotImp}. Throughout this paper we set $f_{\rm
  MS}=1$.

We assume a Salpeter IMF $\mu(\Mstar)\sim M^{-2.35}$, truncated at $0.1
\Msun$ on the low mass end and at $100 \Msun$ on the high mass
end. The corresponding mean stellar mass, $\bar{m}_*$ is 0.35 $\Msun$.

For the turnoff mass, $M_{\rm TO}$, we take the following fitting
formula

\begin{align}
\log(M_{\rm TO})=0.24 + 0.068 x^2-0.34 x+4.76 e^{-4.58 x},
\end{align}

where $x=\log(t/10^9 {\rm years})$ and $M_{\rm TO}$ is in units
$\Msun$. This fit is designed to reproduce the results in 
Table 9 of \citet{MaederMeynet:1987a} and then asymptotes to the fit
for $M_{\rm TO}$ given in equation (9) of \citet{CiottiOstriker:2007a}
for intermediate and late times ($t \gsim 10^8$ years).

For $\Delta M(t)$ we use equation (10) from \citet{CiottiOstriker:2007a}

\begin{align}
\Delta M=
\begin{cases}
0.945 M_{\rm TO}-0.503 & M_{\rm TO} < 9 \Msun\\
 M_{\rm TO}-1.4 \Msun &  M_{\rm TO} \ge 9 \Msun
\end{cases}
\end{align}

To estimate the mass loss from main sequence stars
$\dot{m}(\Mstar, t)$,
we use equation 4 \citet{SchroderCuntz:2005a}\footnote{This
  prescription is a generalization of the Reimers' mass loss law. This
expression is derived assuming the stellar wind results from the
turbulent overflow of moaterial in the chromosphere or underneath it.} {\bf AG:
  unfortunately this is not really meant for MS stars...}

\begin{align}
  \dot{m}(\Mstar)=8 \times 10^{-14} \frac{L_* r_*}{\Mstar}
  \left(\frac{T_{\rm eff}}{\rm 4000 K}\right)^{3.5}
  \left(1+\frac{g_{\odot}}{4300 g_*}\right) \Msun \pyear,\
\end{align}

where  $R_*$, $L_*$, $T_{\rm eff}$ and $g_*$ are the stellar radius,
luminosity, effective and surface gravity respectively. $g_{\odot}$ is
the stellar surface gravity. To calculate $R_*$ and $L_*$ we use the
following scaling relations (taken from Kippenhann and Weygert Figures
22.2 and 22.3).

\begin{align}
L_*=
\begin{cases}
L_{\odot} (\Mstar/\Msun)^{3.2} & \Mstar > \Msun \\
L_{\odot} (\Mstar/\Msun)^{2.5} & \Mstar \le \Msun
\end{cases}
\end{align}

\begin{align}
r_*=
\begin{cases}
R_{\odot} (\Mstar/\Msun)^{0.57} & \Mstar > \Msun \\
R_{\odot} (\Mstar/\Msun)^{0.8} & \Mstar \le \Msun
\end{cases}
\end{align}

@@

\begin{align}
\bar{v}_w(t)=2 \dot{\bar{e}}(t)/\dot{\bar{m}}(t)
\label{eq:vwImp}
\end{align}

@@

>>>>

\subsection{Single Burst}

For a stellar population that forms impulsively of age $t$, the rate of mass and energy injection per unit stellar mass are given, respectively, by 
\begin{align} 
  \dot{\bar{m}}(t) &= \frac{\Delta M(t) \mu|_{M_{\rm TO}(t)}
    \left|\dot{M}_{\rm TO}(t)\right| + f_{\rm MS} \int_{m_0}^{m_{\rm
        T}(t)}
    \dot{m}(\Mstar, t) \mu|_{\Mstar} {\rm d}\Mstar }{\bar{m}_*}\\
  \dot{\bar{e}}(t) &=\dot{e}_{\rm TO}(t)+ f_{\rm MS} \int_{m_0}^{m_{\rm T}(t)}
  \frac{\vw^2(\Mstar, t) \dot{m}(\Mstar, t) \mu|_{\Mstar} {\rm d}\Mstar}{\bar{m}_*},
  \label{eq:edotImp}
\end{align} 
where the first and second terms in each expression corresponds to contributions from main sequence stars and post-main sequence (PMS) stars, respectively.  Here $ f_{\rm MS} < 1$ is an efficiency, which accounts for the efficiency with which main sequence winds thermalize their energy with the rest of the injected gas (we hereafter assume $f_{\rm MS} = 1$) and $\mu$ is the IMF, which we take to be of the Salpeter form $\mu(\Mstar)\sim M^{-2.35}$, truncated at $m_0 = 0.1 \Msun$ on the low mass end and at $100 \Msun$ on the high mass end, where $\bar{m}_*$ = 0.35 $\Msun$ is the corresponding mean stellar mass.  The turn-off mass $M_{\rm TO}$ at time $t$ is calculated according to the following functional fit
\begin{align}
\log(M_{\rm TO})/M_{\odot} =0.24 + 0.068 x^2-0.34 x+4.76 e^{-4.58 x},
\end{align}
where $x=\log(t/10^9 {\rm years})$, so designed to reproduce the results of \citet{MaederMeynet:1987a} (their Table 9) for massive stars while asymptoting to the fitting formula provided by \citet{CiottiOstriker:2007a} (their eq.~[9]) for intermediate and late times ($t \gsim 10^8$ years).

The mass lost in PMS winds $\Delta M(t)$ is calculating from \citet{CiottiOstriker:2007a} (their eq.~[10]) according to
\begin{align}
\Delta M=
\begin{cases}
0.945 M_{\rm TO}-0.503 & M_{\rm TO} < 9 \Msun\\
 M_{\rm TO}-1.4 \Msun &  M_{\rm TO} \ge 9 \Msun
\end{cases}
\end{align}
The MS wind mass loss rate $\dot{m}(\Mstar, t)$ is calculated based on the generalization of Reimer's law given in \citet{SchroderCuntz:2005a} (their eq.~[4]):
\begin{align}
  \dot{m}(\Mstar)=8 \times 10^{-14} \frac{L_* r_*}{\Mstar}
  \left(\frac{T_{\rm eff}}{\rm 4000 K}\right)^{3.5}
  \left(1+\frac{g_{\odot}}{4300 g_*}\right) \Msun \pyear,\
\end{align}
where  $R_*$, $L_*$, $T_{\rm eff}$ and $g_*$ are the stellar radius,
luminosity, effective temperature, and surface gravity, respectively, the latter normalized to its stellar value $g_{\odot}$.  The stellar radius and luminosity are estimated as 
\begin{align}
L_*=
\begin{cases}
L_{\odot} (\Mstar/\Msun)^{3.2} & \Mstar > \Msun \\
L_{\odot} (\Mstar/\Msun)^{2.5} & \Mstar \le \Msun
\end{cases}
\end{align}
\begin{align}
R_*=
\begin{cases}
R_{\odot} (\Mstar/\Msun)^{0.57} & \Mstar > \Msun \\
R_{\odot} (\Mstar/\Msun)^{0.8} & \Mstar \le \Msun
\end{cases}
\end{align}
based on \citealt{Kippenhahn&Weigert} (their Figs.~22.2, 22.3)

@@

\begin{align}
\bar{v}_w(t)=2 \dot{\bar{e}}(t)/\dot{\bar{m}}(t)
\label{eq:vwImp}
\end{align}
while {\bf BDM: insert correct expression}
\begin{align}
\eta = \dot{\bar{m}}(t)
\label{eq:vwImp}
\end{align}

@@

\subsection{Over Star Formation History}
